---
title: "Impact of Tax Liens in New York City: Preliminary Findings"
author: "Fanhua, Sai Priya Kodidala, Raihan Amir"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
warnings: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r library, echo=FALSE,warning=FALSE,message=FALSE}
#install.packages("gridExtra")
#install.packages("grid")
#install.packages("stargazer")

library(tidyverse)
library(fastDummies)
library(ggplot2) #included with Tidyverse
library(forcats) #included with Tidyverse
library(foreign)     #note only imports .dta files up thru Stata 12
library(readstata13)  #imports .dta files from Stata 13 thru 15
library(dplyr)
library(lubridate)
library(weights)
# Install and load the readr package

library(readr)

library(knitr)
library(xtable)
library(readxl)
library(moments)
library(zoo)
library(gridExtra)
library(grid)

library(broom)
library(knitr)

library(stargazer)

```

```{r load data, echo = FALSE}
sales_demo_econ_clean <- readRDS("sales_demo_econ_clean.rds")
load("newdata2014.RData")
```


# 1. Introduction

## <span style="font-size: 24px;"> Problem Statement and Hypothesis </span> 

Every year, homeowners in New York City who fall behind on their tax or water bills can find themselves subject to the City’s annual lien sale, a process through which the City sells outstanding tax and water debts, and the ability to collect them, to private investors.

The majority of these homeowners live in communities of color and are already financially vulnerable

To better understand how the lien sale impacts homeowners and their communities, we conducted an analysis of liens sold on properties by the City from 2013 to 2021

# 2. Background


```{r }

newdata$ovshare <- newdata$Hsp1P + newdata$BlNHP

```


## <span style="font-size: 24px;"> Sales vs Hispanic/Black Neighborhood in 2014 </span> 
```{r echo=FALSE, fig.cap="Sales vs Hispanic/Black Neighborhood"}
ggplot() +
    geom_sf(data = newdata, aes(fill = ovshare), color = "white", size = 0.1) +
    scale_fill_gradient(low = "black", high = "lightblue", limits = c(0, 100)) +
    geom_point(data = filter(newdata, year == 2014, cxy_lon > -75), aes(x = cxy_lon, y = cxy_lat),  color = "red", alpha = 0.8, size = 0.05) +
    labs(title = "Sales vs Hispanic/Black Neighborhood", fill = "% of Hispanic or Black communities") +
    theme_minimal() +
    theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom")
```

## <span style="font-size: 24px;">Data Descriptions </span>

```{r }

colnames(sales_demo_econ_clean)

```

## <span style="font-size: 24px;">Data Descriptions </span>

New York City Tax Lien Sale Data from 2013 to 2021

New York City Demographics and Economic Data from 2021 ACS 5-year survey

Year
```{r }

summary(sales_demo_econ_clean$year)
```
Sales Count
```{r }
summary(sales_demo_econ_clean$sales_count)
```

## <span style="font-size: 24px;">Data Descriptions </span>
% of Black/African American population
```{r }
summary(sales_demo_econ_clean$BlNHP)
```
% of Hispanic/Latino population
```{r }
summary(sales_demo_econ_clean$Hsp1P)
```
Median Income
```{r }
summary(sales_demo_econ_clean$medincom)
```
# Preliminary Findings

## <span style="font-size: 24px;">Black Homeownership and Tax Lien Property Numbers </span>

```{r plot_race1, echo=FALSE, fig.cap="Scatterplot: % Black vs. Tax Lien Sales"}
plot_race1 <- ggplot(data = sales_demo_econ_clean, 
          aes(x = BlNHP, y = sales_count)) + 
     geom_point() +
     geom_smooth(method = 'lm', formula = y ~ x) +
     xlab("Share of black residents") + 
     ylab("Number of properties on tax lien sales") 
```


```{r plot_race2, echo=FALSE, fig.cap="Scatterplot: % Black vs. Tax Lien Sales"}
 plot_race2 <- ggplot(data = sales_demo_econ_clean, 
           aes(x = BlNHP, 
               y = sales_count,
               size = Pop_2E)) + 
      geom_point(alpha = 0.1) + 
      scale_size(range = c(0.1, 6), guide = "none")+  
     xlab("Share of black residents") + 
     ylab("Number of properties on tax lien sales") +
     labs(size = "Total population")
```


```{r plot1, warning=FALSE,message=FALSE}

grid.arrange(plot_race1, plot_race2, nrow = 1)

# Add correlation value
cor_value <- cor(sales_demo_econ_clean$BlNHP, sales_demo_econ_clean$sales_count, use = "pairwise.complete.obs")
cor_text1 <- paste("Correlation: ", round(cor_value, 3))

grid.text(label = cor_text1, x = 0.5, y = -0.2, just = "center", gp = gpar(fontsize = 14))

```
- `r cor_text1`

## <span style="font-size: 24px;">Median Income and Tax Lien Property Numbers </span>

```{r plot_medinc1, echo=FALSE, fig.cap="Scatterplot: % Income vs. Tax Lien Sales"}
 #scatterplot: median income vs tax lien sales
    plot_medinc1 <- ggplot(data = na.omit(sales_demo_econ_clean), 
                           aes(x = medincom, y = sales_count)) + 
      geom_point() +
      geom_smooth(method = 'lm', formula = y ~ x) 
    ggsave("plot_medinc1.png", plot = plot_medinc1, width = 8, height = 6, units = "in")
    
```

```{r plot_medinc2, echo=FALSE, fig.cap="Scatterplot: % Income vs. Tax Lien Sales"}
    plot_medinc2 <- ggplot(data = na.omit(sales_demo_econ_clean), 
                           aes(x = medincom,
                               y = sales_count,
                               size = Pop_2E)) + 
      geom_point(alpha = 0.1) + #alpha adjusts the transparency of points
      scale_size(range = c(0.1, 6), guide = "none") 
    
    ggsave("scatterplot_medinc2.png", plot = plot_medinc2, width = 8, height = 6, units = "in")
```


```{r plot2, warning=FALSE,message=FALSE}

grid.arrange(plot_medinc1, plot_medinc2, nrow = 1)

# Add correlation value
cor_value2 <- cor(sales_demo_econ_clean$medincom, sales_demo_econ_clean$sales_count, use = "pairwise.complete.obs")
cor_text2 <- paste("Correlation: ", round(cor_value2, 3))

grid.text(label = cor_text2, x = 0.5, y = -0.2, just = "center", gp = gpar(fontsize = 14))

```
- `r cor_text2`

## <span style="font-size: 24px;">Median Income & Race and Tax Lien Property Numbers </span>

```{r plot_medinc_race,echo=FALSE}
plot_medinc_race <- ggplot(data = na.omit(sales_demo_econ_clean), 
                               aes(x = BlNHP, 
                                   y = sales_count, 
                                   color = medincom,
                                   size = medincom)) + 
      geom_point(alpha = 0.1) +
      scale_size(range = c(0.1,6), guide = "none") + 
      scale_color_gradient(low = "purple", high ="red") + 
      xlab("Share of black residents") + 
      ylab("Number of tax lien sales") + 
      labs(color = "Median income")
    
    ggsave("scatterplot_medinc&race.png", plot = plot_medinc_race, width = 8, height = 6, units = "in")
    
```

```{r plot3, warning=FALSE}
grid.arrange(plot_medinc_race, nrow = 1)

```

## Preliminary Regression Results
<div style="font-size: 20px;">
\[ sales_t = \beta_0 + \beta_1 black_t + \beta_2 hispanic_t + \beta_3 medincome_t + \epsilon_t \]
</div>

```{r regression_results, echo=FALSE, warning=FALSE, results='asis'}

# Fit the regression models
model1 <- lm(sales_count ~ BlNHP + medincom + factor(year), data = sales_demo_econ_clean)
model3 <- lm(sales_count ~ BlNHP + Hsp1P + medincom + factor(year), data = sales_demo_econ_clean)
model4 <- lm(sales_count ~ BlNHP + Hsp1P + medincom + Borough.y + factor(year), data = sales_demo_econ_clean)

# Generate regression table
table_html <- stargazer::stargazer(model1, model3, model4, type = "html", title = "Regression Results", align = TRUE, header = FALSE, font.size = "small", summary = FALSE)

# Add CSS styling to the table
styled_table <- paste0("<style> table { font-size: 9px; } </style>", table_html)

# Print the styled table
cat(styled_table)

```

